---
title: "Read data"
output:
  html_document:
    highlight: tango
    toc: true
---

## Read in data and explore
```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(readxl)
library(stringr)
library(lubridate)

get_num_list <- function(val){
  # reformat the `val` column, e.g. get_num_list('#3-#5') = '3,4,5'
  if (grepl('-', val) || nrow(str_locate_all(val, '#')[[1]]) > 1){
    return(str_split(val, '-|#') %>% 
      unlist() %>%
      gsub('#|,', '', .) %>% 
      as.integer() %>% 
      na.omit() %>% 
      as.list() %>% 
      do.call('seq.int', .) %>% 
      paste(collapse = ','))
  }
  gsub('#', '', val)
}
```

```{r include=FALSE}
# case_deets <- read_excel('data/cases-detail.xlsx')
# labs <- read_excel('data/labs.xlsx')
# nms <- names(read_excel('data/cases.xlsx', n_max = 0))
# ct <- vector(mode = 'character')
# for (i in nms){
#   ct <- c(ct, if (grepl('^em_yn|premedi_smoking|nurse_|caseid', i)){
#   'text'
#   } else if (grepl('postop_tnt_peak30|preop_tnt', i)) { # very high missing rate for these columns anyway
#     'text'
#   } else {'guess'})
# }
# 
# ecg_patients <- readr::read_csv('data/vitaldb_trks_art_ecg.csv') %>%
#   mutate(Filename = gsub('.vital', '', Filename)) %>%
#   pull(Filename)
# 
# cases <- read_excel('data/cases.xlsx', col_types = ct) %>%
#   mutate(fileid = gsub('.vital', '', fileid)) %>%
#   filter(fileid %in% ecg_patients) %>% # only consider invasive arterial monitoring  patients
#   mutate(em_yn = as.factor(em_yn),
#          fileid = gsub('.vital', '', fileid))
# 
# tf <- read_excel('data/tf.xlsx') %>%
#   mutate(fileid = gsub('.vital', '', fileid)) %>%
#   filter(fileid %in% ecg_patients) %>% # only consider invasive arterial monitoring  patients
#   distinct() # there are some duplicated rows in tf
# skimr::skim(tf)
# 
# save(tf, ecg_patients, cases, file = 'data/gen-outcome.Rdata')
```


```{r warning=FALSE}
load('data/gen-outcome.Rdata')

massive_trans <- tf %>% 
  mutate(val_ori = val) %>% 
  filter(grepl('#', val), # only consider rows with #
         !grepl('cont', val),
         !val %in% c(' #', '#!', '#0316454', 'on #0')) %>% 
  mutate(val = gsub('[?()]', '', val) %>% 
           gsub('[~=]', '-', .) %>% 
           gsub('#9-`1', '#9-12', .) %>% # continuous transfusion for D1_160720_000512
           gsub('#1400cc', '#1', .) %>% # first transfusion of 400cc for D4_191003_082024
           gsub('#17,18,1', '#17,18', .) %>% # continuous transfusion for E5_180907_075903
           gsub('#11,12,1', '#11,12', .) %>% # continuous transfusion for A4_180528_103731
           gsub('#25-289', '#25-28', .) %>% # continuous transfusion for D1_180419_143306
           gsub('#35-#474', '#35-44', .) %>% # continuous transfusion for F4_160803_115114
           gsub('#9#10#11', '#9-11', .) %>% # D5_160901_172629
           gsub('#7,8,9,1', '#7-10', .) %>% # D4_190619_094830
           gsub('#3-1', '#1', .) %>% # C2_190905_073807, C5_190507_120916
           gsub('#3-2', '#2', .) %>% # C2_190905_073807, C5_190507_120916, A2_180716_111059
           gsub('#3-3', '#3', .) %>% # C2_190905_073807, C5_190507_120916, A2_180716_111059
           gsub('on #2,3', NA, .) %>% # E6_180517_124933
           gsub('FMS|FM|CPB|B\\+|hold|on|fro| |resume|pheres', '', .))

```

```{r}
# length(unique(massive_trans$fileid))
# length(unique(tf$fileid))
# setdiff(unique(tf$fileid), unique(massive_trans$fileid))
```


## Preprocessing of specific patients:

The following changes include some straight forward modifications and others (also checked with `sum_rbc`, see Seung Mi's google doc).

- We confirm that `F2_180816_095513` has `sum_rbc= 6`, concluding they did have massive transfusion.
Replacing `#1-3` with `#3-5` and `#3` with `#6`.

- Only one row for `D4_191111_185103`. 
Checked `sum_rbc = 1`. 
So this patient did not have massive transfusion.

- `A2_181126_101740`: `sum_rbc = 13`, had massive transfusion.

- `A2_181126_101740`: `sum_rbc = 6`, had massive transfusion.

- `C4_190508_124038` does not have massive transfusion. 
Checked `sum_rbc = 3`. 

- `A5_180504_073943` (`sum_rbc = 4`) and `C4_190823_074052` (`sum_rbc = 6`) and `D1_190905_201426` (`sum_rbc = 3`) didn't have massive transfusion. (see Seung Mi's google doc)
```{r echo=FALSE}
massive_trans[massive_trans$fileid == 'A1_190215_132805', 'val'] <- c('#1,#2', '#3', '#4', '#5', '#6')
massive_trans[massive_trans$fileid == 'F4_180827_112738', 'val'] <- c('#1', '#2', '#3', '#4')
massive_trans[massive_trans$fileid == 'D1_180727_110744' & massive_trans$val == '#3.4.5', 'val'] <- '#3,4,5'
massive_trans[massive_trans$fileid == 'F4_171219_131836', 'val'] <- c('#1', '#2', '#3', '#4')
massive_trans[massive_trans$fileid == 'D5_190109_124529', 'val'] <- c('#1,2', '#3-6', '#7-10', '#11-14', '#15')
massive_trans[massive_trans$fileid == 'F2_180816_095513', 'val'] <- c('#1', '#2', '#3-5', '#6')
massive_trans[massive_trans$fileid == 'D4_191111_185103', 'val'] <- '#1'
massive_trans[massive_trans$fileid == 'C4_190508_124038', 'val'] <- c('#1', '#2', '#3')
massive_trans[massive_trans$fileid == 'D1_190905_201426', 'val'] <- c('#1-2', '#3')
massive_trans[massive_trans$fileid == 'C4_190823_074052', 'val'] <- paste0('#', seq.int(6))
massive_trans[massive_trans$fileid == 'A5_180504_073943', 'val'] <- c(paste0('#', seq.int(4)), NA)
massive_trans[massive_trans$fileid == 'A2_181126_101740', 'val'] <- c('#1', '#2', '#3-4', '#5-6', '#7-8', paste0('#', 9:13))
massive_trans[massive_trans$fileid == 'D1_170724_102446', 'val'] <- c('#1-2', '#3-4', '#5-6')
massive_trans[massive_trans$fileid == 'D1_190923_095547', 'val'] <- c('#1-4', '#5-6', '#7-8')
massive_trans[massive_trans$fileid == 'D2_190218_171557', 'val'] <- paste0('#', seq.int(5))
massive_trans[massive_trans$fileid == 'D5_170111_105538', 'val'] <- c('#1-11', '#12-27', '#28-43', '#44-51', '#52-59') 
massive_trans[massive_trans$fileid == 'F1_180517_214132', 'val'] <- c('#1-2', '#3-6', '#7-10', '#11-14')
massive_trans[massive_trans$fileid == 'F2_190210_060623', 'val'] <- c("#1-3", "#4-6", "#7-9", "#10-14", "#15")
massive_trans[massive_trans$fileid == 'D5_180803_162219', 'val'] <- c("#1-2", "#3-4", "#5-10", "#11-14", "#15-18", "#19-22", "#23-24")
massive_trans %>% 
  filter(fileid %in% 
           c('A1_190215_132805', 'F4_180827_112738', 'D1_180727_110744', 'D5_190109_124529', 
             'F4_171219_131836', 'F2_180816_095513', 'D4_191111_185103', 'C4_190508_124038', 
             'A5_180504_073943', 'C4_190823_074052', 'D1_190905_201426', 'A2_181126_101740', 
             'D1_170724_102446', 'D1_190923_095547', 'D2_190218_171557', 'D5_170111_105538',
             'F1_180517_214132', 'F2_190210_060623')) %>% 
  DT::datatable()
```

## Systematic analysis

```{r}
# massive_trans %>% filter(grepl('\\.', val)) %>% select(-fileid)
# Also needs to replace . with -
massive_trans <- massive_trans %>% 
  filter(val != '#') %>% 
  mutate(val = gsub('\\.', '-', val))

massive_trans$conv_val <- sapply(massive_trans$val, get_num_list)
```


```{r}
all_trans <- massive_trans %>%
  mutate(sep_val = conv_val) %>% 
  tidyr::separate_rows(sep_val, sep = ',') %>%
  mutate(sep_val = as.integer(sep_val))

patients_to_check <- unique(all_trans$fileid)

pheno_list <- list()

for (patient in patients_to_check){
  df_i <- all_trans %>% filter(fileid == patient)
  start_h <- min(df_i$dt)
  masstf <- 0.5

  for (i in 3:max(df_i$sep_val)){
    if ((i %in% df_i$sep_val) && (sum(df_i$sep_val <= (i-2)) > 0)){
      pre_idx <- max(df_i$sep_val[df_i$sep_val <= (i-2)])

      t2 <- df_i %>% filter(sep_val == i) %>% pull(dt)
      t1 <- df_i %>% filter(sep_val == pre_idx) %>% pull(dt)
      diff_h <- difftime(t2, t1, units = 'hours') %>% as.numeric()
      if (diff_h < 1){
        start_h <- t1[1]
        masstf <- 1
        break
      }
    }
  }
  
  if (!is.na(start_h)) {
    pheno_list[[patient]] <- data.frame(fileid = patient,
                                        masstf = masstf,
                                        correct_start_time = start_h)
  }
}

pheno_df <- bind_rows(pheno_list)
```


```{r}
pheno_df %>%
  mutate(correct_start_time = as_datetime(correct_start_time)) %>% 
  readr::write_csv('data/2021-03-23_auto-mass-trans-phenotype.csv')
```
